Road traffic sound level estimation from realistic urban sound mixtures by Non-negative Matrix Factorization

### 2018-03-13 paper/articleNmfTrafficSimScene2018.tex
minor updates

Several variants of NMF are tested. The threshold initialized NMF appears to be the most relibale approach, allowing road traffic noise level estimation with average errors of 1.3 dB over the tested corpus of sound scenes.

@@

An other approach, followed in this paper, is to consider the blind source separation paradigm which consists in the extraction of a specific signal inside a set of mixed signals, see Figure \ref{fig:source_separation}. From the different existing methods, Non-negative Matrix Factorization (NMF) \cite{lee_learning_1999}, appers to be a relevant method for monophonic sensor networks. Many applications can be found for musical \cite{smaragdis_non-negative_2003, benetos2006musical} and speech \cite{wilson_speech_2008, mysore2011non} contents. Dedicated to sound separation with environmental sound, Immani and Kasa\"i \cite{satoshi_innami_nmf-based_2012} used NMF in a two steps sound separation with the help of time variant gain features. 
A first study \cite{gloaguen2018Estimation} has been conducted, in which diverse NMF estimation rules are compared, namely the supervised, the semi-supervised, and the threshold initialized NMF, have been applied on a large set of simulated sound scenes. This corpus mixes 6 sound categories (\textit{alert, animals, climate, humans, mechanics, transportation}) with a traffic component calibrated to different sound levels to obtain variable traffic predominance. 
The diversity of this corpus was made to assess the performances and the limits of each NMF formula. 
However, if it reveals the interest of NMF for urban sound environments, the assessment of its performance on a corpus of realistic sound scenes must be carried out in order to implement it on a sensor network. 
Design realistic urban sound mixtures makes it possible to access to many acoustic proprieties as the onset and offset time and the sound level of each sound class and especially the traffic component. However, like all simulated process, the issue of its realism has to be resolved in order to validate the NMF performances.

@@

Non-negative Matrix Factorization (NMF) is a linear approximation method proposed by Paatero and Tapper \cite{paatero1994positive} and popularized by Lee and Seung \cite{lee_learning_1999}. It consists in approximating a non negative matrix $\mathbf{V}$ $\in \mathbf{R}^+_{F \times N}$ by the product of two non negative matrices: $\mathbf{W}$, called \textit{dictionary} (or basis), and $\mathbf{H}$, called the matrix \textit{activation} with the dimensions $F \times K$ and $K \times N$ respectively.

@@

The choice of the dimensions is often made such as $F\times K + K \times N < F \times N$ so that NMF can be a low rank approximation. This condition however is not mandatory. When  applying NMF to audio data, $\mathbf{V}$ is usually considered as the magnitude spectrogram obtained by a Short-Time Frequency Transform, $\mathbf{W}$ includes audio spectra and $\mathbf{H}$ is equivalent to the temporal activation of each spectrum, see Figure \ref{fig:exampleNMF}. Because of the non-negativity constraint, only additive combinations between the elements of $\mathbf{W}$ are considered. %The dictionary is then composed of elementary elements providing a part based representation.

@@

The main issue with the supervised approach is the representational limit imposed by a fixed $\mathbf{W}$. To be completely successful, all the acoustical sources must be considered in it which is not possible in an urban environment. To overcome this issue, semi-supervised NMF (SEM-NMF) \cite{lee_semi-supervised_2010} is considered. It consists in decomposing, $\mathbf{W}_{F \times (K+J)}$ into two distinctive matrices: $\mathbf{W} = \left[ \mathbf{W_s}~\mathbf{W_r} \right]$ where $\mathbf{W_s}_{F \times K}$ is a fixed part of $\mathbf{W}$ composed of audio spectra and $ \mathbf{W_r}_{F \times J}$, a mobile part which is updated, see eq. \ref{eq:W_r_SS}. Thus it is possible to include elements not present in $\mathbf{W_s}$. The dimension of $\mathbf{W_r}$ is set up as $J << K$ in order to best considered the sound source present in $\mathbf{W_s}$. $\mathbf{H}$ is then also decomposed in two matrices, $\mathbf{H}_{(K+J) \times N} = \genfrac[]{0pt}{0}{\mathbf{H_s}}{\mathbf{H_r}}$. The eq. \ref{eq:nmf} becomes

@@

{\scriptsize
\begin{subequations}\label{eq:WH-SSupdate}
\begin{align}
\mathbf{W_r}^{(i+1)} &\leftarrow \mathbf{W_r}^{(i)}\otimes\left(\frac{\left[\left(\mathbf{W_r H_r}^{(i)} \right)^{(\beta-2)}\otimes\mathbf{V} \right]\mathbf{H_r}^T}{\left(\mathbf{W_r H_r}^{(i)} \right)^{(\beta-1)}\mathbf{H_r}^T}\right)^{\gamma(\beta)}.\label{eq:W_r_SS}\\
\mathbf{H_r}^{(i+1)} &\leftarrow \mathbf{H_r}^{(i)}\otimes\left(\frac{\mathbf{W_r}^T \left[\left(\mathbf{W_r H_r}^{(i)} \right)^{(\beta-2)}\otimes\mathbf{V} \right]}{\mathbf{W_r}^T \left(\mathbf{W_r H_r}^{(i)} \right)^{(\beta-1)}}\right)^{\gamma(\beta)},\label{eq:H_r_SS}\\
\mathbf{H_s}^{(i+1)} &\leftarrow \mathbf{H_s}^{(i)}\otimes\left(\frac{\mathbf{W_s}^T \left[\left(\mathbf{W_s H_s}^{(i)} \right)^{(\beta-2)}\otimes\mathbf{V} \right]}{\mathbf{W_s}^T \left(\mathbf{W_s H_s}^{(i)} \right)^{(\beta-1)}}\right)^{\gamma(\beta)},\label{eq:H_s_SS}
\end{align}
\end{subequations}}

In this study, $\mathbf{W_s}$ is composed of traffic audio spectra, as it is the sound source of interest, to include in $\mathbf{W_r}$ all the other sources that can be present in the urban sound scenes. The traffic signal estimation is next defined by the fixed part,

@@

where $\mathbf{w}$ is a $k$ element of $\mathbf{W}$ of $F \times 1$ dimension. When $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$=1, both element are identicals. If $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$=0, the element $\mathbf{w'}$ is fully different. The extraction of traffic elements in $\mathbf{W'}$ is carried out by a hard thresholding method \cite{donoho1994threshold}. It consists in weighting in a binary way the traffic elements $\mathbf{W'}$ according to $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$ and a threshold value such as:

\ml{on pourrait tout a fait imaginer l'ajout d'élément de dictionaire random pour se rapprocher de la semi.}

@@

To sum-up, to approximate the audio spectrogram, $\mathbf{V}$ and estimate the traffic component, 3 methods are used. SUP-NMF is only based on the \textit{a priori} knowledge of the traffic to built the dictionary and is constrained to used it to obtain $\mathbf{\tilde{V}}_{traffic}$. SEM-NMF completes this dictionary by the add of a mobile dictionary to consider the interfering sound events and then offer more flexibility. Finally, TI-NMF, as SUP-NMF, considers first a dictionary composed of \textit{traffic} spectra to then update them. Therefore, it is the entire dictionary that can be fully generalized to the sound scenes. The initiated dictionary allows the focused on the source of interest during the updates.

@@

The evaluation corpus is both required to be realistic and non ambiguous in terms of traffic sound level. The former calls for the use of recordings of urban sound environments, and the latter imposes the use of controlled sound sources with known properties: onset and offset time location, and sound level. Indeed, the precise annotation of the sound level of a given source of interest in a polyphonic mixture is not feasible. We thus propose to design such evaluation corpus based on two corpora. The first, called the reference corpus, is a set of recordings of urban sound environments, from which a sort of transcription is made, \textit{i.e.} which sound source is present from this time to this time, at which sound level. The second, called the elemental corpus, is a set of monophonic recordings of sound sources that are occurs in the scenes of the reference. The elemental corpus is built-up with still existing sound and with specific recordings, see part \ref{part:simScene}. The evaluation corpus is then built by replicating recordings of the elemental corpus according to the transcript of the reference corpus, see Figure \ref{fig:transcription}.

More precisely, the urban sound scenes are taken from 76 recordings from 2 to 5 min, achieved in the \nth{13} district of Paris (France) at 19 different locations \footnote{Recordings were made as part of the Grafic project funded by Ademe}, which cover four various sound environments (Figure \ref{fig:map_grafic}), see Figure \ref{fig:map_grafic}. A complete description of the experimental protocol can be found in \cite{aumond_modelling_2017}. Two of the 76 recordings are rejected for the analysis because the audio files were corrupted, resulting in 74 valid audio files assumed as representative of the variety of sound environments. The recordings are listened and categorized within four different sound environments, as proposed in \cite{can_describing_2015}: \textit{park} (P, 8 audio files with a cumulative duration of 16min01), \textit{quiet street} (Q, 33 audio files with a cumulative duration of 77min27), \textit{noisy street} (N, 24 audio files with a cumulative duration of 56min10) and \textit{very noisy street} (vN, 9 audio files with a cumulative duration of 21min42). Then, each audio file is annotated, noticing the start, end time and level of each sound event along with its sound class. This annotation phase makes it possible to obtain simulated sound scenes with the same positions of sound events than the recordings and therefore as close as possible to the realistic scenes.

@@

To transcribe the recordings in simulated scenes, a high quality sound database (wav format, 44.1 kHz sampling rate, high \textit{Signal Noise Ratio}) has been built-up from audio samples found online (\textit{freesound.org}) or with the help of an already existing sound database \cite{salamon2014dataset}. The sound database is composed of two categories of sound: the \textit{event} category, which includes 245 brief sound samples considered as salient, with a 1 to 20 seconds duration and classified among 21 sound classes (\textit{ringing bell, whistling bird, car horn, passing car, hammer, barking dog, siren, footstep, metallic noise, voice \dots}) and the \textit{background} (or \textit{texture}) category gathering 154 long duration sounds ($\approx$ 1m30), whose acoustic properties do not vary in time. This category includes among others\textit{whistling bird, crowd noise, rain, children playing in schoolyard, constant traffic noise} \dots sound classes. Each sound class is composed of multiple samples (\textit{carHorn01.wav, carHorn02.wav \dots}) that are randomly chosen by the \textit{simScene} software to bring diversity.
As the road traffic is the main component in urban environment and is the sound source of interest, recordings of car passages has been made on the Ifsttar's runway. The recordings has been made for 4 cars (Renault Scenic, Renault M\'egane, Renault Clio and Dacia Sandero),  at different speeds and gear ratios. Overall, 103 car passages have been recorded. \ml{ajouter une description sur les problématiques d'overfitting et de generalisation. Cela doit etre un rappel d'un texte ecrit dans la partie nmf. A ce titre, cela pourrait etre une extension de l'etude, on en reparle}. In order to avoid overfitting issues, the audio samples of the first two cars (Renault Scenic and Renault M\'egane) are included in the \textit{SimScene}'s sound database (50 audio files in total). The last 53 audio samples are dedicated to the dictionary design as part of NMF, see section \ref{part:dictionary_building}. A full description of the recordings can be found in \cite{gloaguen_creation_2017}.

@@

With this built-up database, the \textit{SimScene} software and the annotations of the recordings, 74 simulated sound scenes are generated, which have the temporal structure of the reference recordings. The \textit{ebr} parameter is adjusted manually on each sound scene to be faithful compared to the recorded scenes. To validate this adjustment, the mean \textit{Traffic Interfering Ratio} is calculated and summarized on Table \ref{tab:mTIR}. It expresses, on all the scenes of an sound environment, the mean difference between the equivalent traffic sound levels of each scene, $L_{p,traffic}$, with the sound level of the \textit{interfering} sound class, $L_{p, interfering}$, which gathered all the other sound sources not related to the traffic, see Eq \ref{eq:mTIR}. It quantifies the predominance of the traffic component for the 4 types of sound environments.

@@

As it is the sound environment where the traffic is less present, the $mTIR$ is always negative for the \textit{park}. The \textit{interfering} sound class is therefore the main sound source. In the case of the \textit{quiet street}, the $TIR$ can be positive or negative depending on the traffic presence on the scene. For the 2 others sound environments, when the traffic become the main sound source, $mTIR$ is always positive.

@@

The perceptual test is conducted with a panel of 50 subjects that are asked to assess the level of realism on a 7-point scale (1 is \textit{not realistic at all}, 7 is \textit{very realistic}) of a set of transcribed and recorded scenes. The total number of sound scenes tested is set at 40. The first half includes 20 30-seconds audio files, including 5 scenes that belong to the sound environment \textit{park}, 6 from \textit{quiet street}, 4 from \textit{noisy street} and 5 from \textit{very noisy street} chosen randomly among the recorded scenes. The second half is composed of the transcribed scenes coresponding to the 20 reference scenes. In order to limit the duration of the test to preserve the concentration of the subjects, each subject listens to only a subset of 20 sound scenes. All the scenes are normalized to the same sound level, chosen at 65 dB, and the subjects are not allowed to change the output sound level once set at the beginning of the experiment.

The experimental design is elaborated following a partially Balanced Incomplete Block Design (PBIBD) \cite{john1977optimal} that determines the listening order of each participant based on fixed parameters (number of listeners, total number of audio, number of audio assessed by each listeners). The listening order is then built in a such way than each listener listens the same amount of recorded scenes and transcribed scenes to avoide statistical biais. The experimental design and the listening order per participant are performed with the package \textit{sensoMineR} on the \textit{R} software \cite{le_sensominer:_2008}.\\

The test was available online on the 8 February 2017 and the needed number of participant has been reached 12 days later. During the test, the participants had the possibility to listen to each scene as many times as wanted before assessing, without being able to change their judgment afterwards. The participants could also leave a comment on each audio to explain the rating \footnote{The interface of the experiment is available \url{http://soundthings.org/research/xpRealism}}. Based on the information provided, the panel of 50 listeners was made of 31 males and 18 females (one not documented) with an average age of 36 ($\pm$ 12) years old. $62\%$ of the participants declared having no experience in the listening of urban sound mixtures.

A Student's t-test is performed with the $H_0$ hypothesis which considers the similarity between the distribution of the average rates for the recorded and the transcribed scenes of each participant with threshold \textit{p-value} of 5$\%$. Results are displayed in Table \ref{tab:ttest_type_of_scene}.

@@

These results confirm that the recorded and the transcribed scenes are perceived in a similar way by the panel (\textit{p-value} $>$ $\alpha$). More details on the results can be found in  \cite{gloaguen_creation_2017}.

@@

@@

The dictionary is designed from a second sound database specially dedicated to this task to prevent any overfitting issues. It contains  53 audio files of the 2 cars (Dacia Sandero and Renault Clio) recorded on the runway, see part \ref{part:simScene}. Those recordings have not been considered for the creation of the evaluation dataset.

@@

A $\mathbf{K}$-means clustering algorithm is applied to reduce these dimensions to $\mathbf{K} \in \lbrace$25, 50, 100, 200$\rbrace$ in order to avoid redundant information and decrease the computation time. The resulting $\mathbf{K}$ centroids are taken as the elements of $\mathbf{W}$. 
The obtained dictionary is expressed in third octave bands to decrease the dimensionality. Each basis of $\mathbf{W}$ is then normalized as $ \Vert \mathbf{w}_k \Vert = 1$ where $\Vert \bullet \Vert$ is the $\ell_1$ norm. Table \ref{tab:experimental_factorsNMF} summarizes the different modalities of the two experimental factors ($\mathbf{K}$ and $\mathbf{w_t}$).

@@

NMF is performed for 3 $\beta$-divergences: $\beta$ = 2 (euclidean distance), $\beta$ = 1 (Kullback-Leibler divergence) and $\beta = 0$ (Itakura-Sa\"ito divergence). The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are represented in a logarithmic scale through a third band octave representation that reduces the high frequency predominance where the traffic component is absent. In addition, as the number of frequency bins is reduced ($F$ = 29), the computation time is reduced too. 400 iterations are performed to get a stabilized results. For SEM-NMF, the number of elements in $\mathbf{W_r}$ is set to $J = 2$. For hard thresholding, the threshold value, $\mathbf{t_h}$ is set between 0.30 and 0.60 with a 0.01 increment.
Each unique association of modalities between each experimental factor forms an experimental setting. For the filter estimator, 24 settings are computed (4 $\times$ 6). For SUP and SEM-NMF, 240 settings are computed (4 $\times$ 2 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 3). Finally, for TI-NMF, the number of settings is much higher (3720) due to the higher cardinality of the set of threshold values (4 $\times$ 1 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 3 $\times$ 31).
The experimental factors and their different modalities are displayed on Tables \ref{tab:experimental_factorsFilter} and \ref{tab:experimental_factorsNMF}.

@@

\begin{tabularx}{17.5cm}{L{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}C{2cm}@{}}
    \textbf{sound environment} & park & quiet street & noisy street & very noisy street & 4 \\
\end{tabularx}

@@

\begin{tabularx}{17.5cm}{L{3cm}@{}C{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}C{2cm}@{}}
    \textbf{sound environment} & park & quiet street & noisy street & very noisy street & 4\\
\end{tabularx}

>>>>

Several variants of NMF are tested. The threshold initialized NMF appears to be the most reliable approach, allowing road traffic noise level estimation with average errors of less than 1.3 dB over the tested corpus of sound scenes.

@@

An other approach, followed in this paper, is to consider the blind source separation paradigm which consists in the extraction of a specific signal inside a set of mixed signals, see Figure \ref{fig:source_separation}. From the different existing methods, Non-negative Matrix Factorization (NMF) \cite{lee_learning_1999}, appears to be a relevant method for monophonic sensor networks. Many applications can be found for musical \cite{smaragdis_non-negative_2003, benetos2006musical} and speech \cite{wilson_speech_2008, mysore2011non} contents. Dedicated to sound separation with environmental sound, Immani and Kasa\"i \cite{satoshi_innami_nmf-based_2012} used NMF in a two steps sound separation with the help of time variant gain features. Dedicated to the traffic sound separation, a first study \cite{gloaguen2018Estimation} has been conducted, in which diverse NMF estimation rules are compared, namely the supervised, the semi-supervised, and the threshold initialized NMF, have been applied on a large set of simulated sound scenes. This corpus mixes 6 sound categories (\textit{alert, animals, climate, humans, mechanics, transportation}) with a traffic component calibrated to different sound levels, according to the other sound classes (in the rest of the document, these sound classes are resumed as the \textit{interfering} sound class), to obtain variable traffic predominance. 
The diversity of this corpus was made to assess the performances and the limits of each NMF formula. 
However, if this study reveals the interest of NMF for urban sound environments, the assessment of its performance on a corpus of realistic sound scenes must be carried out in order to implement it on a sensor network. 
Design urban sound mixtures makes it possible to access to many acoustic proprieties as the onset and offset time and the sound level of each sound class and especially the traffic component. The realistic aspect of such a corpus is essential to obtain sound scenes similar to recordings and then validate NMF performances. However, like all simulated process, the issue of its realism has to be resolved.

@@

Non-negative Matrix Factorization (NMF) is a linear approximation method proposed by Paatero and Tapper \cite{paatero1994positive} and popularized by Lee and Seung \cite{lee_learning_1999}. It consists in approximating a non negative matrix $\mathbf{V}$ $\in \mathbf{R}^+_{F \times N}$ by the product of two non negative matrices: $\mathbf{W}$, called \textit{dictionary} (or basis), and $\mathbf{H}$, called the \textit{matrix activation} with dimensions $F \times K$ and $K \times N$ respectively,

@@

The choice of the dimensions is often made such as $F\times K + K \times N < F \times N$ so that NMF can be a low rank approximation. This condition however is not mandatory. When  applying NMF to audio data, $\mathbf{V}$ is usually considered as the magnitude spectrogram obtained by a Short-Time Frequency Transform, $\mathbf{W}$ includes audio spectra and $\mathbf{H}$ is equivalent to the temporal activation of each spectrum, see Figure \ref{fig:exampleNMF}. Because of the non-negativity constraint, only additive combinations between the elements of $\mathbf{W}$ are considered.

@@

The main issue with the supervised approach is the representational limit imposed by a fixed $\mathbf{W}$. To be completely successful, all the acoustical sources must be considered in it which is not possible in an urban environment. To overcome this issue, semi-supervised NMF (SEM-NMF) \cite{lee_semi-supervised_2010} has been proposed. It consists in decomposing, $\mathbf{W}_{F \times (K+J)}$ into two distinctive matrices: $\mathbf{W} = \left[ \mathbf{W_s}~\mathbf{W_r} \right]$ where $\mathbf{W_s}_{F \times K}$ is a fixed part of $\mathbf{W}$ composed of audio spectra and $ \mathbf{W_r}_{F \times J}$, a mobile part which is updated, see eq. \ref{eq:W_r_SS}. Thus it is possible to include in it elements not present in $\mathbf{W_s}$. The dimension of $\mathbf{W_r}$ is set up as $J << K$ in order to best considered the sound source present in $\mathbf{W_s}$. $\mathbf{H}$ is then also decomposed in two matrices, $\mathbf{H}_{(K+J) \times N} = \genfrac[]{0pt}{0}{\mathbf{H_s}}{\mathbf{H_r}}$. The eq. \ref{eq:nmf} becomes

@@

{\scriptsize
\begin{subequations}\label{eq:WH-SSupdate}
\begin{align}
\mathbf{W_r}^{(i+1)} &\leftarrow \mathbf{W_r}^{(i)}\otimes\left(\frac{\left[\left(\mathbf{W_r H_r}^{(i)} \right)^{(\beta-2)}\otimes\mathbf{V} \right]\mathbf{H_r}^T}{\left(\mathbf{W_r H_r}^{(i)} \right)^{(\beta-1)}\mathbf{H_r}^T}\right)^{\gamma(\beta)}, \label{eq:W_r_SS}\\
\mathbf{H_r}^{(i+1)} &\leftarrow \mathbf{H_r}^{(i)}\otimes\left(\frac{\mathbf{W_r}^T \left[\left(\mathbf{W_r H_r}^{(i)} \right)^{(\beta-2)}\otimes\mathbf{V} \right]}{\mathbf{W_r}^T \left(\mathbf{W_r H_r}^{(i)} \right)^{(\beta-1)}}\right)^{\gamma(\beta)},\label{eq:H_r_SS}\\
\mathbf{H_s}^{(i+1)} &\leftarrow \mathbf{H_s}^{(i)}\otimes\left(\frac{\mathbf{W_s}^T \left[\left(\mathbf{W_s H_s}^{(i)} \right)^{(\beta-2)}\otimes\mathbf{V} \right]}{\mathbf{W_s}^T \left(\mathbf{W_s H_s}^{(i)} \right)^{(\beta-1)}}\right)^{\gamma(\beta)}.\label{eq:H_s_SS}
\end{align}
\end{subequations}}

In this study, $\mathbf{W_s}$ is composed of traffic audio spectra, as it is the sound source of interest, to include in $\mathbf{W_r}$ the interfering class that can be present in the urban sound scenes. The traffic signal estimation is next defined by the fixed part,

@@

where $\mathbf{w}$ is a $k$ element of $\mathbf{W}$ of $F \times 1$ dimensions. When $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$=1, both element are identicals. If $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$=0, the element $\mathbf{w'}$ is fully different from $\mathbf{w_0}$. The extraction of traffic elements in $\mathbf{W'}$ is carried out by a hard thresholding method \cite{donoho1994threshold}. It consists in weighting in a binary way $\mathbf{W'}$ according to $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$ and a threshold value such as:

%\ml{on pourrait tout a fait imaginer l'ajout d'élément de dictionaire random pour se rapprocher de la semi.}

@@

To summarize, to approximate the audio spectrogram, $\mathbf{V}$ and estimate the traffic component, 3 methods are used which deal differently the \textit{a priori} knowledge on the traffic. SUP-NMF is only based on fixed \textit{traffic} elements and is constrained to used it to obtain $\mathbf{\tilde{V}}_{traffic}$. SEM-NMF completes this dictionary by the add of a mobile dictionary to consider the interfering sound events and then offer more flexibility. Finally, TI-NMF considers first a dictionary composed of \textit{traffic} spectra, as SUP-NMF, but to then update them. This method therefore makes it the most suitable for solving the generalization issues as it is the entire dictionary that can be fully adapted to the sound scenes. The interest of an initiated dictionary is then to focus on the source of interest during the updates.

@@

The evaluation corpus is both required to be realistic and non ambiguous in terms of traffic sound level. The former calls for the use of recordings of urban sound environments, and the latter imposes the use of controlled sound sources with known properties: onset and offset time location, and sound level. Indeed, the precise annotation of the sound level of a given source of interest in a polyphonic mixture is not feasible. Two corpora are proposed to design such a corpus of  evaluation. The first, called the reference corpus, is a set of recordings of urban sound environments, from which a sort of transcription is made, \textit{i.e.} which sound source is present from this time to this time, at which sound level. The second, called the elemental corpus, is a set of monophonic recordings of isolated sound sources that are occurs in the scenes of the reference. The elemental corpus is built with still existing sound and with specific recordings, see part \ref{part:simScene}. The evaluation corpus is then designed by replicating the reference corpus according to the transcription with the help of the elemental corpus, see Figure \ref{fig:transcription}.
More precisely, the urban sound scenes are taken from 76 recordings from 2 to 5 min, achieved in the \nth{13} district of Paris (France) at 19 different locations \footnote{Recordings were made as part of the Grafic project funded by Ademe}, which cover four various sound environments (Figure \ref{fig:map_grafic}), see Figure \ref{fig:map_grafic}. A complete description of the experimental protocol can be found in \cite{aumond_modelling_2017}. Two of the 76 recordings are rejected for the analysis because the audio files were corrupted, resulting in 74 valid audio files assumed as representative of the variety of sound environments. The recordings are listened and categorized within four different sound environments, as proposed in \cite{can_describing_2015}: \textit{park} (P, 8 audio files with a cumulative duration of 16min01), \textit{quiet street} (Q, 35 audio files with a cumulative duration of 77min27), \textit{noisy street} (N, 23 audio files with a cumulative duration of 56min10) and \textit{very noisy street} (vN, 8 audio files with a cumulative duration of 21min42). Then, each audio file is transcripted, noticing the start, end time and level of each sound event along with its sound class. This annotation phase makes it possible to obtain simulated sound scenes with the same positions of sound events than the recordings and therefore as close as possible to the realistic scenes.

@@

To transcribe the recordings in simulated scenes, a high quality sound database (wav format, 44.1 kHz sampling rate, high \textit{Signal Noise Ratio}) has been built-up from audio samples found online (\textit{freesound.org}) or with the help of an already existing sound database \cite{salamon2014dataset}. The sound database is composed of two categories of sound: the \textit{event} category, which includes 245 brief sound samples considered as salient, with a 1 to 20 seconds duration and classified among 21 sound classes (\textit{ringing bell, whistling bird, car horn, passing car, hammer, barking dog, siren, footstep, metallic noise, voice \dots}) and the \textit{background} (or \textit{texture}) category gathering 154 long duration sounds ($\approx$ 1m30), whose acoustic properties do not vary in time. This category includes among others \textit{whistling bird, crowd noise, rain, children playing in schoolyard, constant traffic noise} \dots~sound classes. Each sound class is composed of multiple samples (\textit{carHorn01.wav, carHorn02.wav \dots}) that are randomly chosen by the \textit{simScene} software to bring diversity.
As the road traffic is the main component in urban environment and is the sound source of interest, recordings of car passages has been made on the Ifsttar's runway. The recordings has been made for 4 cars (Renault Scenic, Renault M\'egane, Renault Clio and Dacia Sandero),  at different speeds and gear ratios. Overall, 103 car passages have been recorded. In order to avoid overfitting issues, the audio samples of the first two cars (Renault Scenic and Renault M\'egane) are included in the \textit{SimScene}'s sound database (50 audio files in total). The last 53 audio samples are dedicated to the dictionary design as part of NMF, see section \ref{part:dictionary_building}. A full description of the recordings can be found in \cite{gloaguen_creation_2017}.

% \ml{ajouter une description sur les problématiques d'overfitting et de generalisation. Cela doit etre un rappel d'un texte ecrit dans la partie nmf. A ce titre, cela pourrait etre une extension de l'etude, on en reparle}.

@@

With this built-up database, the \textit{SimScene} software and the transcriptions of the recordings, 74 simulated sound scenes are generated, which have the temporal structure of the reference recordings. The sound level of each sound class is adjusted manually on each sound scene to be faithful compared to the recorded scenes. To validate this adjustment, the mean \textit{Traffic Interfering Ratio} ($mTIR$) is calculated, see Table \ref{tab:mTIR}. It expresses, on all the scenes of an sound environment, the mean difference between the equivalent traffic sound levels of each scene, $L_{p,traffic}$, with the sound level of the \textit{interfering} sound class, $L_{p, interfering}$, which gathered all the other sound sources not related to the traffic, see Eq \ref{eq:mTIR}. It quantifies the predominance of the traffic component for the 4 types of sound environments.

@@

As it is the sound environment where the traffic is less present, the $mTIR$ is always negative for the \textit{park}. The \textit{interfering} sound class is therefore the main sound source which is coherent with this kind of environment. In the case of the \textit{quiet street}, the $TIR$ can be positive or negative depending on the traffic presence on the scene. For the 2 others sound environments, when the traffic become the main sound source, $mTIR$ is always positive.

@@

The perceptual test is conducted with a panel of 50 subjects that are asked to assess the level of realism on a 7-point scale (1 is \textit{not realistic at all}, 7 is \textit{very realistic}) of a set of transcribed and recorded scenes. The total number of sound scenes tested is set at 40. The first half includes 20 30-seconds audio files, including 5 scenes that belong to the sound environment \textit{park}, 6 from \textit{quiet street}, 4 from \textit{noisy street} and 5 from \textit{very noisy street} chosen randomly among the recorded scenes. The second half is composed of the transcribed scenes corresponding to the same 20 reference scenes. In order to limit the duration of the test to preserve the concentration of the subjects, each subject listens to only a subset of 20 sound scenes. All the scenes are normalized to the same sound level, chosen at 65 dB, and the subjects are not allowed to change the output sound level once set at the beginning of the experiment.

The experimental design is elaborated following a partially Balanced Incomplete Block Design (PBIBD) \cite{john1977optimal} that determines the listening order of each participant based on fixed parameters (number of listeners, total number of audio, number of audio assessed by each listeners). The listening order is then built in a such way than each listener listens almost the same amount of recorded scenes and transcribed scenes to avoid statistical biais. The experimental design and the listening order per participant are performed with the package \textit{sensoMineR} on the \textit{R} software \cite{le_sensominer:_2008}.\\

The test was available online on the 8 February 2017 and the needed number of participant has been reached 12 days later. During the test, the participants had the possibility to listen to each scene as many times as wanted before assessing, without being able to change their judgment afterwards. The participants could also leave a comment on each audio to explain the rating \footnote{The interface of the experiment is available \url{http://soundthings.org/research/xpRealism}}. Based on the information provided, the panel of 50 listeners was made of 31 males and 18 females (one not documented) with an average age of 36 ($\pm$ 12) years old. $62\%$ of the participants declared having no experience in the listening of urban sound mixtures. 
The average rate of all the transcribed scenes ($m_{transcribed} = 5.1 (\pm 1.6)$) is then similar to the recorded ones ($m_{recorded} = 4.9 (\pm 1.6)$). In complement, a Student's t-test is performed with the $H_0$ hypothesis which considers the similarity between the distribution of the average rates for the recorded and the transcribed scenes of each participant with a \textit{p-value} threshold of 5$\%$. Results are displayed in Table \ref{tab:ttest_type_of_scene}.

@@

This result confirms that the recorded and the transcribed scenes are perceived in a similar way by the panel (\textit{p-value} $>$ $\alpha$). More details on the results can be found in  \cite{gloaguen_creation_2017}.

@@

The experimental factors related to this estimator are summarized in Table \ref{tab:experimental_factorsFilter}.

@@

The dictionary is designed from a sound database specially dedicated to this task. To prevent any overfitting issues, it contains 53 audio files of the 2 other cars (Dacia Sandero and Renault Clio) recorded on the runway, see part \ref{part:simScene}. Those recordings have not been considered for the creation of the evaluation dataset.

@@

A $\mathbf{K}$-means clustering algorithm is applied to reduce these dimensions to $\mathbf{K} \in \lbrace$25, 50, 100, 200$\rbrace$ in order to avoid redundant information and decrease the computation time. The resulting $\mathbf{K}$ clusters are taken as the elements of $\mathbf{W}$. 
The obtained dictionary is expressed in third octave bands to decrease the dimensionality. Each basis of $\mathbf{W}$ is then normalized as $ \Vert \mathbf{w}_k \Vert = 1$ where $\Vert \bullet \Vert$ is the $\ell_1$ norm. Table \ref{tab:experimental_factorsNMF} summarizes the different modalities of the two experimental factors ($\mathbf{K}$ and $\mathbf{w_t}$).

@@

NMF is performed for 3 $\beta$-divergences: $\beta$ = 2 (euclidean distance), $\beta$ = 1 (Kullback-Leibler divergence) and $\beta = 0$ (Itakura-Sa\"ito divergence). The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are represented in a logarithmic scale through a third band octave representation that reduces the high frequency predominance where the traffic component is absent. In addition, as the number of frequency bins is reduced ($F$ = 29), the computation time is reduced too. 400 iterations are performed to get a stabilized results. For SEM-NMF, the number of elements in $\mathbf{W_r}$ is set to $J = 2$. For hard thresholding, the threshold value, $\mathbf{t_h}$ is set between 0.30 and 0.60 with a 0.01 increment.
Each unique association of modalities between each experimental factor forms an experimental setting. For the filter estimator, 24 settings are computed (4 $\times$ 6). For SUP and SEM-NMF, 192 settings are computed (4 $\times$ 2 $\times$ 2 $\times$ 4 $\times$ 3). Finally, for TI-NMF, the number of settings is much higher (2976) due to the higher cardinality of the set of threshold values (4 $\times$ 1 $\times$ 2 $\times$ 4  $\times$ 3 $\times$ 31).
The experimental factors and their different modalities are displayed on Tables \ref{tab:experimental_factorsFilter} and \ref{tab:experimental_factorsNMF}.

@@

\begin{tabularx}{17.5cm}{L{3cm}@{}C{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}C{2cm}@{}}
    \textbf{sound environment} & \begin{tabular}[c]{@{}c@{}}park\\ 'P'\end{tabular} & \begin{tabular}[c]{@{}c@{}}quiet street\\ 'Q'\end{tabular} & \begin{tabular}[c]{@{}c@{}}noisy street\\ 'N' \end{tabular}& \begin{tabular}[c]{@{}c@{}}very noisy street\\ 'vN'\end{tabular} & 4\\
\end{tabularx}

@@

\begin{tabularx}{17.5cm}{L{3cm}@{}C{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}@{}C{3cm}@{}C{2cm}@{}}
    \textbf{sound environment} & \begin{tabular}[c]{@{}c@{}}park\\ 'P'\end{tabular} & \begin{tabular}[c]{@{}c@{}}quiet street\\ 'Q'\end{tabular} & \begin{tabular}[c]{@{}c@{}}noisy street\\ 'N' \end{tabular}& \begin{tabular}[c]{@{}c@{}}very noisy street\\ 'vN'\end{tabular} & 4\\
\end{tabularx}
